---
title: "R Notebook"
output: html_notebook
---
# Libraries

```{r}
library(tidyverse)
```

# Data

```{r}
here::here()
cakes <- read_csv(here::here("clean_data/tidy_cakes.csv"))
```


# Q1 - Which cake has the most cocoa in it?

One bowl chocolate

```{r}
cakes %>% 
  filter(ingredient == "Cocoa") %>% #Filter for cakes containing cocoa
  arrange(desc(amount)) #Arrange in descending order
```

# Q2 - For sponge cake, how many cups of ingredients are used in total?

3.5 cups

```{r}
cakes %>% 
  #select Sponge cake ingredients that are measured in cups
  filter(cake == "Sponge" & measure == "cup") %>% 
  #Sum the total amount of ingredients measured in cups
  summarise(sum(amount))
```

# Q3 - How many ingredients are measured in teaspoons?

8

```{r}
cakes %>% 
  select(ingredient, measure) %>% 
#Select rows where ingredients are not duplicated and are measured in teaspoons
  filter(!duplicated(ingredient) & measure == "teaspoon") %>% 
#Count the number of rows
  summarise(num_teaspoons = n())

#Second way using the distinct fuction
#cakes %>% 
  #filter(measure == "teaspoon") %>% 
  #distinct(ingredient) %>% 
  #count()
```

# Q4 - Which cake has the most unique ingredients?

Angel

```{r}
#Find unique ingredients and assign to new object
unique_ingredients <- cakes %>% 
#group cakes by ingredient
  group_by(ingredient) %>% 
#count number of times each ingredient is used
  summarise(num_times_used = n()) %>% 
#filter so that only ingredients used once are included
  filter(num_times_used == 1)

#Join unique ingredients table with cakes table to find cakes with unique ingredients
inner_join(cakes, unique_ingredients, "ingredient") %>% 
#group by cake
  group_by(cake) %>% 
#count the number of unique ingredients each cake contains
  summarise(num_unique_ingredients = n())
```

# Q5 - Which ingredients are used only once?
Bananas, cream of tartar, crushed ice, dried currants, egg white, nutmeg, nuts, zwiebach

```{r}
#Call the unique_ingredients table
unique_ingredients
```

# Q6 - Mean number of ingredients in a single cake

```{r}
cakes %>% 
  select(cake, ingredient) %>% 
  group_by(cake) %>% 
  summarise(num_ingredients = n()) %>% 
  summarise(mean(num_ingredients))
```
