---
title: "R Notebook"
output: html_notebook
---

# 1. Introduction to the dataset

The dataset contains information about cake ingredients for a variety of cakes, including the type of measure and the amount of each ingredient. The dataset is in long format and contains the columns "cake", "ingredient_code", "ingredient", "measure", and "amount".

# 2. List of assumptions

1) Assume that cakes that had NA in the amount column of an ingredient didn't contain that ingredient
2) That NAs in the amount column for the ingredient "Sour cream cup" were measured in cups.
3) That "one" in the measure column means one unit of whatever ingredient is listed because this is present in ingredients like "banana" and "egg".

# 3. Steps taken to clean the data

Note: The code chunks below are for explanation only and won't work, use the 1_clean_cakes_data.R file.

## 1) Convert data to long form

Data needed to be pivoted longer so that ingredients were in one column and amounts were in a separate column.

```{r}
# Pivot data longer - move ingredients to a column and amounts to a separate column
cake_ingredients_long <- cake_ingredients %>% 
  pivot_longer(
    cols = AE:ZH,
    names_to = "ingredient_code",
    values_to = "amount"
  )
```
## 2) Join tables

Only ingredient codes were in the main dataset, so I joined the ingredient code dataset so that the clean dataset would include ingredient names and the type of measure for each. 

```{r}
# Join with cake codes to get ingredient names and type of measure for each
cakes_joined <- cake_ingredients_long %>% 
  left_join(cake_codes, by = c("ingredient_code" = "code")) %>% 
  janitor::clean_names()
```

## 3) Find and deal with missing values

There were two main columns that contained missing values: "amount" and "measure". I assumed that NAs in "amount" meant that the cake didn't contain that ingredient, so these rows were dropped

```{r}
cakes_joined <- cakes_joined  %>% 
  filter(!is.na(amount))
```

I explored NAs in the measure section and found that these were only present for the ingredient "Sour cream cup", so I assumed that this ingredient was measured in cups, but that the information had been mistakenly recorded in the ingredient column.

```{r}
#Replaces NAs in measure section with "cup"
cakes_imputed <- cakes_joined %>% 
  mutate(measure = coalesce(measure, "cup"))
```

# 4. Answers to task brief questions

## 1) Which cake has the most cocoa in it?

The one bowl chocolate cake has the most cocoa in it.

## 2) For sponge cake, how many cups of ingredients are used in total?

For sponge cake, there are 3 ingredients that are measured in cups: sifted flour, granulated sugar and water. When summed up, 3.5 cups of ingredients are used in total.

## 3) How many ingredients are measured in teaspoons?

8 unique ingredients are measured in teaspoons

## 4) Which cake has the most unique ingredients?

Note: Assuming that this question means to find the cake that uses the highest number of ingredients that are not used in any other cake.

The angel cake has the most unique ingredients

## 5) Which ingredients are used only once?

Bananas, cream of tartar, crushed ice, dried currants, egg white, nutmeg, nuts, and  zwiebach are used only once

# 5. Additional analysis

## 1) What is the mean number of ingredients in a cake?
8.67