---
title: "R Notebook"
output: html_notebook
---

# Libraries

```{r}
library(tidyverse)
```

# Read data

```{r}
here::here()

candy_clean <- read_csv(here::here("clean_data/candy_clean.csv"))

```

# Analysis questions

## Q1) What is the total number of candy ratings given across the three years. 
(Number of candy ratings, not the number of raters. Don’t count missing values)

772,352 candy ratings in total

```{r}
candy_clean %>% 
  #selects ratings
  select(rating) %>% 
  #drops nas so that they're not included in the count
  drop_na(rating) %>% 
  #counts the number of ratings
  summarise(num_ratings = n())
  
```

## Q2) What was the average age of people who are going out trick or treating?

Mean age of trick-or-treaters is 35

```{r}
candy_clean %>% 
  select(age, going_out) %>% 
  #filters data to only include those going out
  filter(going_out == "Yes") %>% 
  #drops NAs in the age variable so that mean can be calculated
  drop_na(age) %>% 
  #computes the mean age
  summarise(mean_age_going_out = mean(age))
```


## Q3) What was the average age of people who are not going trick or treating?

Mean age of non trick-or-treaters is 39

```{r}
candy_clean %>% 
  select(age, going_out) %>% 
  #filters data to only those not going out
  filter(going_out == "No") %>% 
  #drops NAs in age so mean can be calculated
  drop_na(age) %>% 
  #calculates the mean age
  summarise(mean_age_staying_in = mean(age))
```


## Q4) For each of joy, despair and meh, which candy bar received the most of these ratings?

Broken glow stick received the most "despair" ratings
Any full-sized candy bar received the most "joy" ratings
Lollipops received the most "meh" ratings

```{r}
candy_clean %>% 
  select(candy, rating) %>% 
  #groups by ratings and then type of candy
  group_by(rating, candy) %>% 
  #counts the number of ratings for each rating category for each candy
  summarise(num_ratings = n()) %>% 
  #takes the highest rated candy for each rating.
  slice_max(order_by = num_ratings, n = 1)
```


## Q5) How many people rated Starburst as despair?

1990 people rated Starburst as despair.

```{r}
candy_clean %>% 
  select(candy, rating) %>% 
  #filters data to only include Starburst and "despair" ratings
  filter(candy == "Starburst" & rating == "DESPAIR") %>% 
  #counts the number of rows(aka despair ratings)
  summarise(num_despair = n())
```


## Add numeric rating variable for next questions

```{r}
candy_clean <- candy_clean %>% 
  #create a new column for a numeric version of rating
  mutate(rating_numeric = case_when(
    #sets values for despair, joy and meh ratings
    rating == "DESPAIR" ~ -1,
    rating == "JOY" ~ 1,
    rating == "MEH" ~ 0
    #NAs will reach the end and not be assigned a numeric value
  )
  )
```

## Q6) What was the most popular candy bar by this rating system for each gender in the dataset ?

For participants of all genders (male, female and unknown), Any full-sized candy bar was the most popular candy bar.

When full-sized candy bars are discounted:
-Female participants prefer cash
-Male participants prefer Reese's peanut butter cups
-Unknown gender participants prefer cash

Since cash is not a candy bar, female and unknown gender participants were analysed further and after cash, they preferred:
-Female participants - Reese's peanut butter cups
-Unknown gender participants - KitKats

```{r}
candy_clean %>% 
  select(gender, candy, rating_numeric) %>% 
  #drop rows where rating is NA so that they can be summed
  drop_na(rating_numeric) %>% 
  #groups by gender and then candy
  group_by(gender, candy) %>% 
  #sums the numeric rating for each candy by gender
  summarise(sum_rating = sum(rating_numeric)) %>% 
  #takes the top 3 candys and their ratings for each gender
  slice_max(order_by = sum_rating, n = 3)
```



## Q7) What was the most popular candy bar in each year?

For every year, Any full-sized candy bar was the most popular candy bar

When full sized candy bars are discounted:
- Reese's peanut butter cups were most popular in 2015
- Cash was most popular in 2016
- Reese's peanut butter cups were most popular in 2017

Since cash is not a candy bar, 2016 was examined further and KitKat was the most popular.

```{r}
candy_clean %>% 
  select(year, candy, rating_numeric) %>% 
  #drops NAs in numeric rating so sum can be calculated
  drop_na(rating_numeric) %>% 
  #groups by year and then candy
  group_by(year, candy) %>% 
  #sums the numeric ratings for each candy for every year
  summarise(sum_rating = sum(rating_numeric)) %>% 
  #takes the top three candies and their ratings for each year
  slice_max(order_by = sum_rating, n = 3)
```

## Q8) What was the most popular candy bar by this rating for people in US, Canada, UK, and all other countries?

For Canada, the USA and all other countries, Any full-sized candy bar was the most popular candy bar. For the UK, cash was the most popular item. The most popular candy bar was a four-way tie between:
-any full-sized candy bar
-Lindt truffles
-Rolo
-Toblerone

When full-sized candy bars are discounted:
-Americans prefer Reese's peanut butter cups
-Canadians prefer KitKats
-People from any other country prefer cash.

Since cash is not a candy bar, "other country" was further examined and KitKat was the most popular

```{r}
candy_clean %>% 
  select(country, candy, rating_numeric) %>% 
  #drops NAs in numeric rating so sum can be calculated
  drop_na(rating_numeric) %>% 
  #groups by country and then candy
  group_by(country, candy) %>% 
  #takes the sum of numeric ratings for each candy for each country
  summarise(sum_rating = sum(rating_numeric)) %>% 
  #takes the top 3 candies and their ratings for each year.
  slice_max(order_by = sum_rating, n = 3)
```



