---
title: "R Notebook"
output: html_notebook
---

# Libraries

```{r}
library(tidyverse)
```

# Read data

```{r}
here::here()

candy_clean <- read_csv(here::here("clean_data/candy_clean.csv"))

```

# Analysis questions

## Q1) What is the total number of candy ratings given across the three years. 
(Number of candy ratings, not the number of raters. Don’t count missing values)

772,352 candy ratings in total

```{r}
candy_clean %>% 
  select(rating) %>% 
  drop_na(rating) %>% 
  summarise(num_ratings = n())
  
```

## Q2) What was the average age of people who are going out trick or treating?

Mean age of trick-or-treaters is 35

```{r}
candy_clean %>% 
  select(age, going_out) %>% 
  filter(going_out == "Yes") %>% 
  drop_na(age) %>% 
  summarise(mean_age_going_out = mean(age))
```


## Q3) What was the average age of people who are not going trick or treating?

Mean age of non trick-or-treaters is 39

```{r}
candy_clean %>% 
  select(age, going_out) %>% 
  filter(going_out == "No") %>% 
  drop_na(age) %>% 
  summarise(mean_age_staying_in = mean(age))
```


## Q4) For each of joy, despair and meh, which candy bar received the most of these ratings?

Broken glow stick received the most "despair" ratings
Any full-sized candy bar received the most "joy" ratings
Lollipops received the most "meh" ratings

```{r}
candy_clean %>% 
  select(candy, rating) %>% 
  group_by(rating, candy) %>% 
  summarise(num_ratings = n()) %>% 
  arrange(rating, desc(num_ratings), candy) %>% 
  slice_max(order_by = num_ratings, n = 1)
```


## Q5) How many people rated Starburst as despair?

1990 people rated Starburst as despair.

```{r}
candy_clean %>% 
  select(candy, rating) %>% 
  filter(candy == "Starburst" & rating == "DESPAIR") %>% 
  summarise(num_despair = n())
```


## Add numeric rating variable for next questions

```{r}
candy_clean <- candy_clean %>% 
  mutate(rating_numeric = case_when(
    rating == "DESPAIR" ~ -1,
    rating == "JOY" ~ 1,
    rating == "MEH" ~ 0
  )
  )
```

## Q6) What was the most popular candy bar by this rating system for each gender in the dataset ?

For participants of all genders (male, female and unknown), Any full-sized candy bar was the most popular candy bar.

When full-sized candy bars are discounted:
-Female participants prefer cash
-Male participants prefer Reese's peanut butter cups
-Unknown gender participants prefer cash

Since cash is not a candy bar, female and unknown gender participants were analysed further and after cash, they preferred:
-Female participants - Reese's peanut butter cups
-Unknown gender participants - KitKats

```{r}
candy_clean %>% 
  select(gender, candy, rating_numeric) %>% 
  drop_na(rating_numeric) %>% 
  group_by(gender, candy) %>% 
  summarise(sum_rating = sum(rating_numeric)) %>% 
  slice_max(order_by = sum_rating, n = 3)
```



## Q7) What was the most popular candy bar in each year?

For every year, Any full-sized candy bar was the most popular candy bar

When full sized candy bars are discounted:
- Reese's peanut butter cups were most popular in 2015
- Cash was most popular in 2016
- Reese's peanut butter cups were most popular in 2017

Since cash is not a candy bar, 2016 was examined further and KitKat was the most popular.

```{r}
candy_clean %>% 
  select(year, candy, rating_numeric) %>% 
  drop_na(rating_numeric) %>% 
  group_by(year, candy) %>% 
  summarise(sum_rating = sum(rating_numeric)) %>% 
  slice_max(order_by = sum_rating, n = 3)
```

## Q8) What was the most popular candy bar by this rating for people in US, Canada, UK, and all other countries?

For Canada, the USA and all other countries, Any full-sized candy bar was the most popular candy bar. For the UK, cash was the most popular item. The most popular candy bar was a four-way tie between:
-any full-sized candy bar
-Lindt truffles
-Rolo
-Toblerone

When full-sized candy bars are discounted:
-Americans prefer Reese's peanut butter cups
-Canadians prefer KitKats
-People from any other country prefer cash.

Since cash is not a candy bar, "other country" was further examined and KitKat was the most popular

```{r}
candy_clean %>% 
  select(country, candy, rating_numeric) %>% 
  drop_na(rating_numeric) %>% 
  group_by(country, candy) %>% 
  summarise(sum_rating = sum(rating_numeric)) %>% 
  slice_max(order_by = sum_rating, n = 3)
```



